---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# msoc

<!-- badges: start -->

[![R-CMD-check](https://github.com/JanMarvin/msoc/workflows/R-CMD-check/badge.svg)](https://github.com/JanMarvin/msoc/actions)
[![r-universe](https://janmarvin.r-universe.dev/badges/msoc)](https://janmarvin.r-universe.dev/msoc)
<!-- badges: end -->

This R package provides the cryptographic library [msoc](https://github.com/herumi/msoffice) and allows encrypting and decrypting office open xml files with R.

## Installation

You can install the development version of `msoc` from
[GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("JanMarvin/msoc")
```

Or from [r-universe](https://r-universe.dev/) with:

``` r
# Enable repository from janmarvin
options(repos = c(
  janmarvin = 'https://janmarvin.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
# Download and install msoc in R
install.packages('msoc')
```

## Introduction

A word of warning. I have no background in cryptography and I am somewhat unable
to maintain the bundled msoc library. Assume that I only made it work and if it
ever breaks, you are on your own. Due to these reasons the package will never
be released to CRAN either.

Before you encrypt a file, you should memorize the password. If you forget the 
password, there is no way to recover the contents of the file.

```{r}
library(msoc)
file <- system.file("extdata", "Untitled1.xlsx", package = "msoc")

# out is an encrypted file and you will 
# be unable to open it without the password
out <- encrypt(file, pass = "msoc")

# out is an decrypted file and readable
# without the password
out <- decrypt(out, pass = "msoc")
```

## An example with [`openxlsx2`](https://github.com/JanMarvin/openxlsx2)

```{r}
library(openxlsx2)
library(msoc)

xlsx <- temp_xlsx()

# let us write some worksheet
wb_workbook()$add_worksheet()$add_data(x = mtcars)$save(xlsx)

# now we can encrypt it
encrypt(xlsx, xlsx, pass = "msoc")

# the file is encrypted, we can not read it
try(wb <- wb_load(xlsx))

# we have to decrypt it first
decrypt(xlsx, xlsx, pass = "msoc")

# now we can load it again
wb_load(xlsx)$to_df() %>% head()
```

## License

This package is licensed under the BSD 3-Clause license and is based on
[`msoffice`](https://github.com/herumi/msoffice) (by Cybozu Labs, Inc;
COPYRIGHT 2007-2015 and Shigeo Mitsunari;COPYRIGHT 2015-2023) and [`cybozulib`](https://github.com/herumi/cybozulib) (by Cybozu Labs, Inc;
COPYRIGHT 2007-2012 and Shigeo Mitsunari; COPYRIGHT 2012-2023). Both release
under the BSD 3-Clause License. 
